---
title: "A7 - Clustering"
author: "Aditya, Vineet"
date: "November 3, 2017"
output: html_document
---

 
## Subproblem 1: Clustering
## Using Spark, perform hierarchical agglomerative clustering and k-means clustering as follows:
>
  + Fuzzy loudness: cluster songs into quiet, medium, and loud
  + Fuzzy length: cluster songs into short, medium, and long
  + Fuzzy tempo: cluster songs into slow, medium, and fast
  + Fuzzy hotness: cluster songs into cool, mild, and hot based on song hotness
  + Combined hotness: cluster songs into cool, mild, and hot based on two dimensions: artist hotness, and song hotness
  
## K-Means Clustering - Local Runs
```{r echo=FALSE,message=FALSE}
duration <- read.csv("outputForDuration/part-00000")
colnames(duration) = c("SongId","Song_Length","Cluster")
library(easyGgplot2)
ggplot2.stripchart(data=duration, main="K Means Cluster on Length of the", xName='Cluster',yName='Song_Length',groupName = 'Cluster',
                xlab="Clusters 0 - Short, 1- Medium , 2 - Long",ylab="Song Length in Seconds",
                colour="red")
```

```{r echo=FALSE,message=FALSE}
tempo <- read.csv("outputFortempo/part-00000")
colnames(tempo) = c("SongId","Song_Tempo","Cluster")
ggplot2.stripchart(data=tempo, main="K Means Cluster on Length of the", xName='Cluster',yName='Song_Tempo',groupName = 'Cluster')

